TCP Congestion Algorithms In Datacenters \\ {\large Large Systems $\cdot$ Project Report}

### 2017-12-24 paper/paper.tex
Paper: formatting, minor fixes, and release 1.0

\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage[hyperref=true, natbib=true, style=numeric, backend=bibtex]{biblatex}
\usepackage{graphicx}
\bibliography{references.bib}

@@

\title{TCP Congestion Algorithms In Datacenters \\
	\vspace{0.3cm}
	{\large LS Project}
}
\date{\today{}}
\author{Luc Gommans, Rick van Gorp}

@@

\section{Abstract}

@@

\section{Introduction}

@@

This report includes a performance and behaviour analysis of different TCP congestion control algorithms in a (simulated) multitenant datacenter network. Mario Hock et al. describe in a paper that BBR shows unfair behaviour compared to the CUBIC congestion control algorithm \cite{bbr-congestion-comparison}. Therefore the fairness between TCP congestion control algorithms will also be included in the analysis. Based on the results, we will recommend datacenters on measures to take to keep providing a reliable environment for customers when multiple congestion control algorithms are active in the network, for example using Quality of Service or discouraging certain congestion control algorithms. The main research question is: {\it How can a datacenter provide a reliable and performing environment for customers when multiple TCP congestion control algorithms are used in the
network?}

In the next section, we will discuss related work. Section \ref{sec:method}
describes our methods: how we selected algorithms to test, how we tested their
performance, and how we chose network conditions to simulate. Section
\ref{sec:results} describes the results of our research: the algorithms we
chose to test are described in \ref{sub:chosen-algos}, a theoretical analysis
of those is described in \ref{sub:algos-theoretical}, and experimental results
can be found in \ref{sub:experiment-results}. We discuss our results in section
\ref{sec:discussion} and draw conclusions in section \ref{sec:conclusion}.

@@

During this research, performance comparison tests of various TCP congestion algorithms will be performed. The most common TCP congestion algorithms will be determined through desk research. The tests will be conducted in an isolated environment. Based on the results of those tests, recommendations will be given related to the measures a datacenter has to take to keep the traffic equally allocated over customers.

@@

In order to select the most common TCP congestion algorithms desk research will be performed. This desk research will show which algorithms are used more often than other algorithms by looking for the default algorithms in Operating Systems. Based on those results a list of TCP congestion algorithms will be composed.

@@

We will create a bottleneck on the path to the user to see how different
congestion algorithms handle this. The connection will be full duplex and have
symmetrical speeds, to avoid any congestion in sending acknowledgements.
Because the server is sending the large volume of data and the client is
responding with the much smaller acknowledgements, the server will be the one
applying the congestion control. Therefore we vary the algorithm settings of
the server and not the client. The client will always be a GNU/Linux machine
using CUBIC, but we assume (due to the aforementioned reasons) that this does
not matter.

@@

We use three physical hosts as servers: two running Ubuntu 17.04 and one
running Microsoft Windows 10. In each experiment, only two hosts will be used.
The Windows system is used because there is no working CTCP
implementation---the default algorithm in Windows since Vista---available for a
modern Linux kernel.

@@

We will test the TCP congestion algorithm for fairness in sharing bandwidth
with other TCP connections using other congestion algorithms. In order to test
the fairness, the bandwidth usage per second per algorithm is measured. If two
different algorithms use similar amounts of bandwidth, they can be said to be
fair relative to each other.

The tests conducted with the algorithms have two environment variables: packet loss and delay. Those environment variables are simulated using \texttt{netem}, which is a network emulation functionality in several Linux distributions \cite{linux-netem}. \texttt{netem} is applied to the incoming side of the interfaces of the clients to emulate the same packet loss and delay. Windows, which is tested as a server, does not support \texttt{netem}. Therefore the packet loss and delay are applied to the incoming interfaces of the clients. Packet loss and delay are chosen as the behaviour of most TCP algorithms depends on those variables \cite{bbr-congestion}\cite{dctcp-congestion}\cite{cubic-tcp-congestion}\cite{compound-tcp-congestion}\cite{bic-tcp-congestion}. The values for the environment variables are chosen based on measurements performed by Verizon \cite{verizon-latency}.

@@

The data set of Verizon \cite{verizon-latency} is used to calculate the packet loss and delay used in this research. The data set includes global information and shows averages per month over the year 2017. The packet loss is calculated by taking the average of the Verizon Business Packet Delivery Statistics for Country Specific Metrics for each country. Those averages were inverted by subtracting the packet delivery percentage from 100, to show the actual packet loss in percentages and not the packet delivery. Based on the ascending sorted values a bar chart was created. This bar chart is shown in figure \ref{fig:packet-loss-chart}

@@

The country numbers in figure \ref{fig:packet-loss-chart} are only used to show the distribution of the average packet loss. The maximum packet loss shown in figure \ref{fig:packet-loss-chart} equals 0.6\% and the minimum equals 0\%. However, in case of situations where more packet loss occurs than usual, the double value of the maximum is used: 1.2\%. The values for packet loss in this comparison set-up were determined based on the distribution shown in figure \ref{fig:packet-loss-chart} and the manually set maximum of 1.2\%. The values are shown in table \ref{table:test-packetloss}.

\begin{table}[H]
	\centering
	\caption{Packet loss percentages used for the comparison setup}
	\begin{tabular}[H]{ | l |}
	\hline
	\textbf{Packet loss} \\
	\hline 0\%. \\
	\hline 0.01\% \\
	\hline 0.1\% \\
	\hline 0.6\% \\
	\hline 1.2\% \\
	\hline
	\end{tabular}
	\label{table:test-packetloss}
\end{table}

The delay is calculated by taking the average of the Verizon Business Latency Statistics for Country Specific Metrics for each country. Based on the ascending sorted values a bar chart was created. This bar chart is shown in figure \ref{fig:verizon-delay-chart}

@@

\begin{table}[H]
	\centering
	\caption{Packet loss percentages used for the comparison setup}
	\begin{tabular}[H]{ | l |}
	\hline
	\textbf{Delay} \\
	\hline  8ms\\
	\hline  64ms  \\
	\hline  120ms  \\
	\hline  176ms \\
	\hline	232ms \\
	\hline  290ms  \\
	\hline
	\end{tabular}
	\label{table:test-delay}
\end{table}

@@

The most common TCP congestion algorithms are CTCP, CUBIC and BIC. CTCP is the default congestion algorithm in Microsoft Windows Server 2008 and newer \cite{cubic-kernel-version}. CUBIC is the default congestion algorithm in Linux distributions with kernel versions 2.6.19 and higher \cite{cubic-kernel-version}. BIC was the default congestion algorithm in earlier Linux distributions from kernel version 2.6.8 until 2.6.19 \cite{bic-kernel-version} \cite{cubic-kernel-version}. DCTCP and BBR will also be tested as those protocols are new and upcoming---they are included in Linux since 4.9 (December 2016) \cite{linux-bbr}. DCTCP is specifically designed for datacenters \cite{dctcp-congestion} and BBR is a new algorithm meant for general use \cite{bbr-congestion}.

@@

The BBR algorithm depends on parameters $BtlBw$ and $RT_{prop}$ \cite{bbr-congestion}. Those parameters are estimates defined by the BBR algorithm. $BtlBw$ specifies the bandwidth at the slowest link in each direction and $RT_{prop}$ is the round-trip propagation time. $RT_{prop} = RTT - queuing delay - processing delay$, which results in the minimum amount of time for round trip propagation in case there are no queuing or processing delays. Using those two parameters, the Bandwidth Delay Product (BDP) is calculated, which is the maximum possible amount of data being sent in a network: $BDP = BtlBw * RT_{prop}$. As long as the amount of data in transit is less than the $BDP$ and $BtlBw$, the delivery rate is increased. If the amount of data in transit is equal to $BtlBw$ the delivery rate can not go up anymore. The $RTT$ can never be lower than $RT_{prop}$... explain more on delay perhaps?.

The CTCP algorithm is a combination of a loss-based and delay-based congestion protocol \cite{compound-tcp-congestion}. For the loss-based component, the variable conventional congestion window, $cwnd$, was introduced. For the delay-based component, the variable delay window, $dwnd$, was introduced. The TCP sending window is determined from: $win = min(cwnd + dwnd, awnd)$, where $awnd$ is the window advertised by the receiver. For $cwnd$ on the arrival of an ACK: $cwnd = cwnd + 1/win$. The $dwnd$ variable is initially set to zero and is only set when the congestion avoidance is active. $dwnd$ is calculated differently for several scenarios. If the network path is underutilised: $dwnd(t+1) = dwnd(t) + \alpha*dwnd(t)^{k} - 1$. When the network path is congested: $dwnd(t+1) = dwnd(t) - \eta*diff$. When packet loss occurs: $dwnd(t+1) = dwnd(t)(1-\beta) - cwnd/2$. $k$, $\alpha$, $\eta$ and $\beta$ are tunable parameters for optimisation of the algorithm. $\eta$ defines the rapidness of reduction of the window when congestion is detected. $\beta$ is the factor used for the multiplicative decrease after loss is detected. $diff$ is calculated from the expected throughput and the actual throughput, which are again calculated from the window size and Round Trip Time (RTT).

The DCTCP algorithm consists of three components \cite{dctcp-congestion-original}:

@@

The BIC algorithm controls its windows based on the size of windows \cite{bic-tcp-congestion}. It consists of two components:

@@

The CUBIC algorithm is a successor of the BIC algorithm \cite{cubic-tcp-congestion}. In case of a loss event $W_{max}$, which is the maximum window size, is registered. This is followed by a multiplicative decrease of the congestion window by factor $\beta$. If an ACK in congestion avoidance is received and the current window size $cwnd$ is less than $W_{max}$, the algorithm uses the concave profile to increase the window size according to the formula: $cwnd = \frac{W(t+RTT)-cwnd}{cwnd}$. If $cwnd$ is larger than $W_{max}$, the convex profile is used to increase the window size according to the same formula as the concave profile. In both formulas $t$ is the time elapsed from the last window reduction.

@@

\begin{figure}[H]
	\makebox[\textwidth]{
		\centering
		\begin{minipage}{0.5\paperwidth}
			\centering
			\includegraphics[width=0.9\textwidth]{figs/bic_vs_cubic_8_0.png}
			\caption{CUBIC vs BIC with 8ms latency and 0\% loss}
			\label{fig:fair-1}
		\end{minipage}\hfill
		\begin{minipage}{0.5\paperwidth}
			\centering
			\includegraphics[width=0.9\textwidth]{figs/dctcp-bic-8-0.png}
			\caption{DCTCP vs BIC with 8ms latency and 0\% loss}
			\label{fig:fair-2}
		\end{minipage}
	}
\end{figure}

@@

\begin{figure}[H]
	\centering
		\includegraphics[scale=0.4]{figs/ctcp-bic-8-0.png}
		\caption{CTCP vs. BIC with 8ms latency and 0\% loss}
	\label{fig:ctcp-1}
\end{figure}

@@

BBR is unique in that it barely responds to packet loss but only to delay: even
with 1.2\% loss, as long as the delay is low (8 milliseconds in our case), it
reaches 850$\pm$20 megabits per second. A comparison between BBR and CUBIC is shown in figure \ref{fig:bbr-cubic}. We see in figure \ref{fig:bbr-cubic} that, compared to CUBIC, the bandwidth usage by the BBR data transfer is higher when 1.2\% packet loss is introduced. However, the link speed of 1Gbps is not fully utilized as is in the paper written by Mario Hock et al. \cite{bbr-congestion-comparison}. We also do not see the CUBIC data transfer responding to the BBR data transfer by reducing its speed in any way. In another test, where CUBIC is compared to BIC with the same parameters, we see CUBIC utilizing the same amount of bandwidth as in the BBR-CUBIC comparison. This is shown in figure \ref{fig:bic-cubic}. In figure \ref{fig:bbr-cubic-0} we show a comparison of a CUBIC data transfer with a BBR data transfer with parameters 8ms latency and 0\% packet loss. This time the total link speed is utilized and we see no unfairness between CUBIC and BBR. We therefore see no signs in our results of BBR being the cause of the CUBIC data transfer being slower. In our results other algorithms also reach much lower
speeds (24 megabits is not uncommon), but we see no signs of BBR being the
cause of that. In tests with BBR nor CTCP, speeds are not higher than when BBR
is one of the algorithms being tested.

@@

\begin{figure}[H]
	\centering
		\includegraphics[scale=0.4]{figs/cubic_vs_bbr_8_0.png}
		\caption{CUBIC vs. BBR with 8ms latency and 0\% loss}
	\label{fig:bbr-cubic-0}
\end{figure}

@@

@@

With the use of a hypothetical TCP congestion control algorithm that is unfair towards other flows, the bandwidth allocation to one customer would be higher than to other customers. When a TCP flow shows an aggressive increase and keeps pushing data over the network, other (fair) algorithms will reduce the transmission rate as packet loss occurs. One important condition for this unfair congestion control algorithm is that it should not be aggressive to the extent of causing a denial of service.

To mitigate this, the bandwidth usage per server could be monitored, prioritising the traffic sent by some servers. The solution uses traffic shaping to control the priority of the traffic per server:

\begin{enumerate}
	\item The monitoring solution measures periodically whether the bandwidth in the link is fully utilised. If this is the case, the next step will be performed. If not, it continues monitoring the link utilisation.
	\item The monitoring solution calculates periodically per server whether its bandwidth usage deviates more than $\alpha$ from the average bandwidth usage. $\alpha$ is deviated bandwidth and is set according to an acceptable deviation value. The following formula is used to calculate the standard deviation: $s_{i} = \sqrt{\frac{\sum_{i=1}^N (x_i - \overline{x})^2}{N-1} }$. $N$ is the amount of active servers on the link. $\overline{x}$ is the average amount of bandwidth for all active servers on the link. $x_i$ is the bandwidth of the current server and $s_{i}$ shows the standard deviation in bandwidth for server $i$. If $s_{i}$ is a negative value, no priority change will be applied as the bandwidth usage is lower than average for the server. If $s_{i} <= \alpha$, then no priority change will be applied. If $s_{i} > \alpha$, the priority will be changed to a lower value.
	\item The monitoring solution provides information of the servers where the priority will be set to a lower value to the traffic shaping controller.
	\item The traffic shaping controller sets the priority for the servers to a lower value, so that the traffic coming from the servers that use $\alpha$ more than average bandwidth has less preference than traffic closer to the average bandwidth and lower.
	\item The monitoring solution periodically checks if the bandwidth in the link is still fully utilised. If this is not the case, the traffic shaping policy is reversed again by setting the priority to the default value. If this is the case, the periodical calculations of server bandwidth usage are continued.
\end{enumerate}

@@

Peng Yang et al. describe a method to remotely identify used TCP congestion
avoidance algorithms\cite{tcp-congestion-identification}. The paper states that
they were able to identify all default TCP algorithms, but some non-default
algorithms could not be identified. The tests performed are limited to remote
web servers, so it requires more research to reliably implement such a method
in a generic manner. If this method is deemed reliable, it can be useful to
detect whether an unfair TCP congestion control algorithm is used as it will
not match any of the default TCP algorithms used. Based on that the traffic of
the unacceptable TCP congestion control algorithms can be shaped by lowering
their priority.

@@

Without any concurrent flows, Windows does reach gigabit speed. On different
hardware, it showed the same, recessive behaviour. We think this might be
caused by the way Python socket operations are handled in Windows. When
repeating the test with \texttt{iperf}, which is written in C, this behaviour
no longer occurs. Even when using the Python equivalent of
\texttt{while(true)\{ socket.send(data); \}}, the bandwidth allocation did not
change. We theorise that the difference is caused by something in the
\texttt{send} call. An alternative theory is that Microsoft Windows for
workstations have a different implementation than Microsoft Windows Server
systems.

BBR is the best-performing algorithm in our set, because it barely responds to
packet loss. However, we were unable to draw the same conclusions to our results as the conclusions written in the paper by Mario Hock et al \cite{bbr-congestion-comparison}. We used multiple values to simulate packet loss as well as delay, which simulate having small buffers. BBR is a TCP congestion control protocol that reacts to the minimum RTT in combination with the Bottleneck Bandwidth. We have tested multiple delay values that appear in the real world. The delay values are directly related to the minimum RTT possible on the link. As mentioned in the results, we see no hard evidence that this
causes other algorithms to reduce their speed further, but there are slight
indications. For example in the test of DCTCP versus BBR, DCTCP reaches around
0.2 to 0.6 megabytes per second. In the test of DCTCP versus BIC, this is 0.3
to 0.7 megabytes per second. The difference is large nor statistically
significant, but it warrants looking into. On the scale of 0.7 megabytes per
second, a tenth of a megabyte does make a difference.

@@

Running a mixed environment of Linux and Windows where Python is running
introduced issues. We expect the send call of Python in Windows is implemented
differently than in Linux, but this requires more testing. The tests for CTCP
should be performed again when the reason for the discrepancy is known and can
be mitigated. The results should also be verified using a Microsoft Windows
Server instance.

@@

BBR appears to perform better in a congested network than the other tested protocols. In order to validate that BBR does not push away other TCP congestion algorithms a new testing method can be used. The data transfer with a TCP congestion algorithm is started and halfway the data transfer with BBR is started. This method shows whether a BBR data transfer has a large influence on the data transfer of other TCP congestion algorithms.

@@

In our results, BBR shows better performance than any other algorithm when
packet loss occurs. Finally, DCTCP is only suitable when both the server and
client are in the same datacenter. There is a considerable ramp-up time when
congestion occurs. Within datacenters this works well, because with no
congestion it keeps the queue lengths to a minimum.

>>>>

\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage[hidelinks]{hyperref}
\usepackage[hyperref=true, natbib=true, style=numeric, backend=bibtex]{biblatex}
\usepackage{graphicx}
\bibliography{references.bib}
\usepackage[top=3.5cm, bottom=3.5cm, left=4.3cm, right=4.3cm]{geometry}

@@

\title{TCP Congestion Algorithms In Datacenters \\
	{\large Large Systems $\cdot$ Project Report}
}
\date{\today{}}
\author{Luc Gommans, Rick van Gorp}

@@

\section*{Abstract}

@@

\newpage
\section{Introduction}

@@

This report includes a performance and behaviour analysis of different TCP
congestion control algorithms in a (simulated) multitenant datacenter network.
Mario Hock et al. describe in a paper that BBR shows unfair behaviour when
compared to the CUBIC congestion control
algorithm\cite{bbr-congestion-comparison}. Therefore the fairness between
congestion control algorithms will also be included in the analysis. Based on
the results, we will recommend datacenters on measures to take to keep
providing a reliable environment for customers when multiple congestion control
algorithms are active in the network, for example using Quality of Service or
prohibiting certain TCP congestion control algorithms. The main research question
is: {\it How can a datacenter provide a reliable and performing environment for
customers when multiple TCP congestion control algorithms are used in the
network?}

In the next section, we will discuss related work. Section \ref{sec:method}
describes our methods: how we selected algorithms to test, how we tested their
performance, and how we chose network conditions to simulate. Section
\ref{sec:results} describes the results of our research: the algorithms we
chose to test are described in \ref{sub:chosen-algos}, a theoretical analysis
of those is described in \ref{sub:algos-theoretical}, and experimental results
can be found in \ref{sub:experiment-results}. We discuss our results in section
\ref{sec:discussion}, mention possible future work in section
\ref{sec:futurework}, and finally draw conclusions in section \ref{sec:conclusion}.

@@

During this research, performance comparison tests of various TCP congestion
control algorithms will be performed. The most common TCP congestion algorithms will be
determined through desk research. The tests will be conducted in an isolated
environment. Based on the results of those tests, recommendations will be given
related to the measures a datacenter has to take to keep the traffic equally
allocated over customers.

@@

In order to select the most common TCP congestion algorithms desk research will
be performed. This desk research will show which algorithms are used more often
than other algorithms by looking for the default algorithms in Operating
Systems. Based on those results, a list of TCP congestion algorithms will be
composed.

@@

We will create a bottleneck on the path to the user to see how different
congestion control algorithms handle this. The connection will be full-duplex
and have symmetrical speeds, to avoid any congestion in sending
acknowledgements. Because the server is sending the large volume of data and
the client is responding with the much smaller acknowledgements, the server
will be the one applying congestion control. Therefore we vary the algorithm
settings of the server and not the client. The client will always be a
GNU/Linux machine using CUBIC, but we assume (due to the aforementioned
reasons) that this does not matter.

@@

We use three physical hosts as servers: two running Ubuntu 17.04 and one
running Microsoft Windows 10. In each experiment, only two hosts will be active.
The Windows system is used because there is no working CTCP
implementation---the default algorithm in Windows since Vista---available for a
modern Linux kernel.

@@

We will test the congestion algorithm for fairness in sharing bandwidth
with other TCP connections using other congestion algorithms. In order to test
the fairness, the bandwidth usage per second per algorithm is measured. If two
different algorithms use similar amounts of bandwidth, they can be said to be
fair relative to each other.

The tests conducted with the algorithms have two environment variables: packet loss and delay. Those environment variables are simulated using \texttt{netem}, which is a network emulation functionality in several Linux distributions\cite{linux-netem}. \texttt{netem} is applied to the incoming side of the interfaces of the clients to emulate the same packet loss and delay. Windows, which is tested as a server, does not support \texttt{netem}. Therefore the packet loss and delay are applied to the incoming interfaces of the clients. Packet loss and delay are chosen as the behaviour of most TCP algorithms depends on those variables\cite{bbr-congestion}\cite{dctcp-congestion}\cite{cubic-tcp-congestion}\cite{compound-tcp-congestion}\cite{bic-tcp-congestion}. The values for the environment variables are chosen based on measurements performed by Verizon\cite{verizon-latency}.

@@

The data set of Verizon\cite{verizon-latency} is used to calculate the packet loss and delay used in this research. The data set includes global information and shows averages per month over the year 2017. The packet loss is calculated by taking the average of the Verizon Business Packet Delivery Statistics for Country Specific Metrics for each country. Those averages were inverted by subtracting the packet delivery percentage from 100, to show the actual packet loss in percentages and not the packet delivery. Based on the ascending sorted values a bar chart was created. This bar chart is shown in figure \ref{fig:packet-loss-chart}

@@

The country numbers in figure \ref{fig:packet-loss-chart} are only used to show
the distribution of the average packet loss. The maximum average packet loss
shown in figure \ref{fig:packet-loss-chart} equals 0.6\% and the minimum equals
0\%. However, to account for situations where more packet loss occurs than usual,
the double value of the maximum is used: 1.2\%. The values for packet loss in
this comparison set-up were determined based on the distribution shown in
figure \ref{fig:packet-loss-chart} and the manually set maximum of 1.2\%. The
values are shown in table \ref{table:test-packetloss}.

\begin{table}[H]
	\centering
	\begin{tabular}[H]{ | l |}
	\hline
	\textbf{Packet loss} \\
	\hline 0\%. \\
	\hline 0.01\% \\
	\hline 0.1\% \\
	\hline 0.6\% \\
	\hline 1.2\% \\
	\hline
	\end{tabular}
	\caption{Packet loss percentages used for the comparison setup}
	\label{table:test-packetloss}
\end{table}

\vspace{0.9cm}

The delay is calculated by taking the average of the Verizon Business Latency
Statistics for Country Specific Metrics for each country. Based on the
ascending sorted values a bar chart was created. This chart is shown in figure
\ref{fig:verizon-delay-chart}.

@@

\begin{table}[H]
	\centering
	\begin{tabular}[H]{ | l |}
	\hline
	\textbf{Delay} \\
	\hline  8ms\\
	\hline  64ms  \\
	\hline  120ms  \\
	\hline  176ms \\
	\hline	232ms \\
	\hline  290ms  \\
	\hline
	\end{tabular}
	\caption{Packet loss percentages used for the comparison setup}
	\label{table:test-delay}
\end{table}

@@

The most common TCP congestion algorithms are CTCP, CUBIC and BIC. CTCP is the default congestion algorithm in Microsoft Windows Server 2008 and newer\cite{cubic-kernel-version}. CUBIC is the default congestion algorithm in Linux distributions with kernel versions 2.6.19 and higher\cite{cubic-kernel-version}. BIC was the default congestion algorithm in earlier Linux distributions from kernel version 2.6.8 until 2.6.19\cite{bic-kernel-version}\cite{cubic-kernel-version}. DCTCP and BBR will also be tested as those protocols are new and upcoming---they are included in Linux since 4.9 (December 2016)\cite{linux-bbr}. DCTCP is specifically designed for datacenters\cite{dctcp-congestion} and BBR is a new algorithm meant for general use\cite{bbr-congestion}.

@@

The BBR algorithm depends on parameters $BtlBw$ and $RT_{prop}$\cite{bbr-congestion}. Those parameters are estimates defined by the BBR algorithm. $BtlBw$ specifies the bandwidth at the slowest link in each direction and $RT_{prop}$ is the round-trip propagation time. $RT_{prop} = RTT - queuing delay - processing delay$, which results in the minimum amount of time for round trip propagation in case there are no queuing or processing delays. Using those two parameters, the Bandwidth Delay Product (BDP) is calculated, which is the maximum possible amount of data being sent in a network: $BDP = BtlBw * RT_{prop}$. As long as the amount of data in transit is less than the $BDP$ and $BtlBw$, the delivery rate is increased. If the amount of data in transit is equal to $BtlBw$ the delivery rate can not go up anymore. The $RTT$ can never be lower than $RT_{prop}$.

The CTCP algorithm is a combination of a loss-based and delay-based congestion protocol\cite{compound-tcp-congestion}. For the loss-based component, the variable conventional congestion window, $cwnd$, was introduced. For the delay-based component, the variable delay window, $dwnd$, was introduced. The TCP sending window is determined from: $win = min(cwnd + dwnd, awnd)$, where $awnd$ is the window advertised by the receiver. For $cwnd$ on the arrival of an ACK: $cwnd = cwnd + 1/win$. The $dwnd$ variable is initially set to zero and is only set when the congestion avoidance is active. $dwnd$ is calculated differently for several scenarios. If the network path is underutilised: $dwnd(t+1) = dwnd(t) + \alpha*dwnd(t)^{k} - 1$. When the network path is congested: $dwnd(t+1) = dwnd(t) - \eta*diff$. When packet loss occurs: $dwnd(t+1) = dwnd(t)(1-\beta) - cwnd/2$. $k$, $\alpha$, $\eta$ and $\beta$ are tunable parameters for optimisation of the algorithm. $\eta$ defines the rapidness of reduction of the window when congestion is detected. $\beta$ is the factor used for the multiplicative decrease after loss is detected. $diff$ is calculated from the expected throughput and the actual throughput, which are again calculated from the window size and Round Trip Time (RTT).

The DCTCP algorithm\cite{dctcp-congestion-original} consists of three components:

@@

The BIC algorithm controls its windows based on the size of windows\cite{bic-tcp-congestion}. It consists of two components:

@@

The CUBIC algorithm is a successor of the BIC algorithm\cite{cubic-tcp-congestion}. In case of a loss event $W_{max}$, which is the maximum window size, is registered. This is followed by a multiplicative decrease of the congestion window by factor $\beta$. If an ACK in congestion avoidance is received and the current window size $cwnd$ is less than $W_{max}$, the algorithm uses the concave profile to increase the window size according to the formula: $cwnd = \frac{W(t+RTT)-cwnd}{cwnd}$. If $cwnd$ is larger than $W_{max}$, the convex profile is used to increase the window size according to the same formula as the concave profile. In both formulas $t$ is the time elapsed from the last window reduction.

@@

\begin{figure}[H]
	\makebox[\textwidth]{
		\centering
		\begin{minipage}{0.5\paperwidth}
			\centering
			\includegraphics[width=0.9\textwidth]{figs/cubic_vs_bic_8_0.png}
			\caption{CUBIC vs BIC with 8ms latency and 0\% loss}
			\label{fig:fair-1}
		\end{minipage}\hfill
		\begin{minipage}{0.5\paperwidth}
			\centering
			\includegraphics[width=0.9\textwidth]{figs/dctcp-bic-8-0.png}
			\caption{DCTCP vs BIC with 8ms latency and 0\% loss}
			\label{fig:fair-2}
		\end{minipage}
	}
\end{figure}

@@

\begin{figure}[H]
	\centering
		\includegraphics[scale=0.55]{figs/ctcp-bic-8-0.png}
		\caption{CTCP vs. BIC with 8ms latency and 0\% loss}
	\label{fig:ctcp-1}
\end{figure}

@@

%\newgeometry{top=1cm,bottom=1.5cm}

BBR is unique in that it barely responds to packet loss but only to delay: even
with 1.2\% loss, as long as the delay is low (8 milliseconds in our case), it
reaches 850$\pm$20 megabits per second. A comparison between BBR and CUBIC is shown in figure \ref{fig:bbr-cubic}. We see in figure \ref{fig:bbr-cubic} that, compared to CUBIC, the bandwidth usage by the BBR data transfer is higher when 1.2\% packet loss is introduced. However, the link speed of 1Gbps is not fully utilized as is in the paper written by Mario Hock et al.\cite{bbr-congestion-comparison}. We also do not see the CUBIC data transfer responding to the BBR data transfer by reducing its speed in any way. In another test, where CUBIC is compared to BIC with the same parameters, we see CUBIC utilizing the same amount of bandwidth as in the BBR-CUBIC comparison. This is shown in figure \ref{fig:bic-cubic}. In figure \ref{fig:bbr-cubic-0} we show a comparison of a CUBIC data transfer with a BBR data transfer with parameters 8ms latency and 0\% packet loss. This time the total link speed is utilized and we see no unfairness between CUBIC and BBR. We therefore see no signs in our results of BBR being the cause of the CUBIC data transfer being slower. In our results other algorithms also reach much lower
speeds (24 megabits is not uncommon), but we see no signs of BBR being the
cause of that. In tests with BBR nor CTCP, speeds are not higher than when BBR
is one of the algorithms being tested.

@@

\begin{figure}[H]
	\centering
		\includegraphics[scale=0.6]{figs/cubic_vs_bbr_8_0.png}
		\caption{CUBIC vs. BBR with 8ms latency and 0\% loss}
	\label{fig:bbr-cubic-0}
\end{figure}

@@

%\restoregeometry

@@

With the use of a hypothetical congestion control algorithm which is unfair towards other flows, the bandwidth allocation to one customer would be higher than to other customers. When a TCP flow shows an aggressive increase and keeps pushing data over the network, other (fair) algorithms will reduce the transmission rate as packet loss occurs. One important condition for this unfair congestion control algorithm is that it should not be aggressive to the extent of causing a denial of service.

To mitigate this, the bandwidth usage per server could be monitored, prioritising the traffic sent by some servers. The monitor uses traffic shaping to control the priority of the traffic per server:

\begin{enumerate}
	\item The monitor measures periodically whether the bandwidth in the link is fully utilised. If this is the case, the next step will be performed. If not, it continues monitoring the link utilisation.
	\item The monitor calculates periodically, per server, whether its bandwidth usage deviates more than $\alpha$ from the average bandwidth usage. $\alpha$ is the deviated bandwidth and is set according to an acceptable deviation value. The following formula is used to calculate the standard deviation: $s_{i} = \sqrt{\frac{\sum_{i=1}^N (x_i - \overline{x})^2}{N-1} }$. $N$ is the amount of active servers on the link. $\overline{x}$ is the average amount of bandwidth for all active servers on the link. $x_i$ is the bandwidth of the current server and $s_{i}$ shows the standard deviation in bandwidth for server $i$. If $s_{i}$ is a negative value, no priority change will be applied as the bandwidth usage is lower than average for the server. If $s_{i} <= \alpha$, then no priority change will be applied. If $s_{i} > \alpha$, the priority will be changed to a lower value.
	\item The monitor provides information of the servers where the priority will be set to a lower value to the traffic shaping controller.
	\item The traffic shaping controller sets the priority for the servers to a lower value, so that the traffic coming from the servers that use $\alpha$ more than average bandwidth has less preference than traffic closer to the average bandwidth and lower.
	\item The monitor periodically checks if the bandwidth in the link is still fully utilised. If this is not the case, the traffic shaping policy is reversed again by setting the priority to the default value. If this is the case, the periodical calculations of server bandwidth usage are continued.
\end{enumerate}

@@

Peng Yang et al.\cite{tcp-congestion-identification} describe a method to
remotely identify used TCP congestion avoidance algorithms. The paper states
that they were able to identify all default TCP algorithms, but some
non-default algorithms could not be identified. The tests performed are limited
to remote web servers, so it requires adaption to implement such a method in a
generic manner. If this method is deemed reliable, it can be useful to detect
whether an unfair algorithm is used as it will not match
any of the default algorithms used. Based on that, the traffic of the
unacceptable TCP congestion control algorithms can be shaped by lowering their
priority.

@@

Without any concurrent flows, Windows does reach the full link speed (1Gbps).
On different hardware, it showed the same, recessive behaviour. We
think this might be caused by the way Python socket operations are handled in
Windows. When repeating the test with \texttt{iperf}, which is written in C,
this behaviour no longer occurs. Even when using the Python equivalent of
\texttt{while(true)\{ socket.send(data); \}}, the bandwidth allocation did not
change. We theorise that the difference is caused by something in the
\texttt{send} call. An alternative theory is that Microsoft Windows for
workstations have a different implementation than Microsoft Windows Server
systems.

BBR is the best-performing algorithm in our set, because it barely responds to
packet loss. However, we were unable to draw the same conclusions to our
results as the conclusions written in the paper by Mario Hock et
al\cite{bbr-congestion-comparison}. We used multiple values to simulate packet
loss as well as delay, which simulate having small buffers. BBR is a TCP
congestion control protocol that reacts to the minimum RTT in combination with
the Bottleneck Bandwidth. We have tested multiple delay values that appear in
the real world. The delay values are directly related to the minimum RTT
possible on the link. As mentioned in the results, we see no hard evidence that
this causes other algorithms to reduce their speed further, but there are
slight indications. For example in the test of DCTCP versus BBR, DCTCP reaches
around 0.2 to 0.6 megabytes per second. In the test of DCTCP versus BIC, this
is 0.3 to 0.7 megabytes per second. The difference is large nor statistically
significant, but it warrants looking into. On the scale of 0.7 megabytes per
second, a tenth of a megabyte does make a difference.

@@

Running a mixed environment of Linux and Windows introduced issues, at least in
combination with Python as benchmark system. We expect that in Windows, the
Python send call on sockets is implemented differently than in Linux, but this
requires more testing. The tests for CTCP should be performed again when the
reason for the discrepancy is known and can be mitigated. The results should
also be verified using a Microsoft Windows Server instance.

@@

BBR performs better in a congested network than any other tested algorithm.
While we did not see clear signs of other algorithms responding to BBR's higher
bandwidth share, this should be tested further to be certain. We suggest
starting a transfer with one algorithm under the desired network circumstances,
determine its bandwidth, and then start a parallel flow using BBR. If the
bandwidth of the original flow does not decrease, BBR is indeed fair towards
this algorithm.

@@

In our results, BBR shows better performance than any other algorithm when
packet loss occurs. Finally, DCTCP is only suitable when both the server and
client are in the same datacenter: there is a considerable ramp-up time when
congestion occurs. Within datacenters DCTCP works well, because without
congestion it keeps queue lengths to a minimum.
